### 2017-03-19 paper/test.tex
Found minted!

\documentclass{article}
\usepackage{minted}
\usepackage{xcolor}
\usemintedstyle{vs}
\definecolor{bg}{rgb}{0.95,0.95,0.95}
\begin{document}
\begin{minted}[mathescape,
               linenos,
               numbersep=5pt,
               tabsize=4,
               label=Helloworld,
               bgcolor=bg,
               breaklines]{TypeScript}
interface Result {
   cand: number,
   votes: number
}
// Calculates the winners in a First-past-the-post election
export default function calculateWinners(input: number[][], seats: number) {

   // Maps input to first prefrences
   let firstPrefrences = input.map(el => {
       return el[0]
   })

   let results: Result[] = []
   // Loops over prefrences and adds/increments result
   for (var i = 0; i < firstPrefrences.length; i++) {
       // Finds index in results array of candidate at i
       let index = results.findIndex(obj => obj.cand === firstPrefrences[i])
       // If candidate is not added to results
       if (index === -1) {
           results.push({ cand: firstPrefrences[i], votes: 1 })
       } else {
           // Increment count if already added in results
           results[index].votes += 1;
       }
   }

   // Sorts array by number of votes
   results.sort((a, b) => {
       if (a.votes > b.votes) {
           return -1
       } else if (a.votes < b.votes) {
           return 1
       }
       return 0
   })

   // Gets the first elements in the sorted array i.e the winners
   let winners = results.slice(0, seats)

   // Returns array of winners
   return winners.map(cand => cand.cand)
}

\end{minted}
\end{document}