### 2017-03-19 paper/test.tex
Minted

\documentclass{article}
\usepackage{minted}
\usepackage{xcolor}
\usemintedstyle{vs}

\renewcommand{\listingscaption}{Code example}
\newcommand{\codeblock}[3]{%
\definecolor{bg}{rgb}{0.95,0.95,0.95}

\begin{listing}[H]
  \inputminted[
    mathescape,
    linenos,
    numbersep=5pt,
    tabsize=4,
    label=Helloworld,
    bgcolor=bg,
    breaklines
  ]{TypeScript}{#1}
  \caption{#3}
  \label{#2}
\end{listing}
}

\begin{document}
\codeblock{code/candidateNode.ts}{code: global data structure}{This is an example}
\end{document}