### 2017-05-22 paper/images/tree.tex
Final touches on the paper

\begin{tikzpicture}[]

\node [align=left,circle, draw](A) at (0,0) {candidate: 0 \\ votes: 38};
\node [align=left,circle, draw](B) at (3.5,0) {candidate: 1 \\ votes: 17};
\node [align=left,circle, draw](C) at (7,0) {candidate: 2 \\ votes: 22};

\node [align=left,circle, draw,] (Aa) at (-1.6,-2.5) {\footnotesize candidate: 1 \\ \footnotesize votes: 11};
\node [align=left,circle, draw,] (Ab) at (0.1,-2.6) { };
\node [align=left,circle, draw,] (Ac) at (0.7,-2.6) { };
\node [align=left,circle, draw,] (Ad) at (1.3,-2.6) { };

\node [align=left,circle, draw,] (Ba) at (2.6,-2.6) { };
\node [align=left,circle, draw,] (Bb) at (3.2,-2.6) { };
\node [align=left,circle, draw,] (Bc) at (3.8,-2.6) { };
\node [align=left,circle, draw,] (Bd) at (4.4,-2.6) { };

\node [align=left,circle, draw,] (Ca) at (6.1,-2.6) { };
\node [align=left,circle, draw,] (Cb) at (6.7,-2.6) { };
\node [align=left,circle, draw,] (Cc) at (7.3,-2.6) { };
\node [align=left,circle, draw,] (Cd) at (7.9,-2.6) { };

\node [align=left,circle, draw,] (Aaa) at (-1,-4.5) { };
\node [align=left,circle, draw,] (Aab) at (-1.6,-4.5) { };
\node [align=left,circle, draw,] (Aac) at (-2.2,-4.5) { };

\draw (A) -- (Aa);
\draw (A) -- (Ab);
\draw (A) -- (Ac);
\draw (A) -- (Ad);

\draw (B) -- (Ba);
\draw (B) -- (Bb);
\draw (B) -- (Bc);
\draw (B) -- (Bd);

\draw (C) -- (Ca);
\draw (C) -- (Cb);
\draw (C) -- (Cc);
\draw (C) -- (Cd);

\draw (Aa) -- (Aaa);
\draw (Aa) -- (Aab);
\draw (Aa) -- (Aac);

\end{tikzpicture}